---
title: "a3_task4_shayan_kaveh"
author: "Shayan Kaveh"
date: "10/24/2021"
output: html_document
---
## Set Up
```{r setup, include= TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(lubridate) # create date classes
library(janitor)
library(here)
library(kableExtra) # make kable table
```

## Overview

This code explores data on red and purple sea urchins recorded by Santa Barbara Coastal Long Term Ecological Research program divers. I wrangle the data to compare the mean sizes of purple urchins at the Naples Reef in 2010 between sites where giant kelp was removed and control sites. 

**Data:** Reed, D. 2020. SBC LTER: Reef: Long-term experiment: Kelp removal: Urchin size frequency distribution ver 20. Environmental Data Initiative. <https://doi.org/10.6073/pasta/5a1e9ef03aa47bd2225c0bb98a02a63b>.

## Read in data

```{r}
# read in urchins data and clean to lowersnake case
urchins <- read_csv(here("data", "urchins.csv")) %>% 
  clean_names()
```

## Compare the difference in mean size of purple urchins at the Naples Reef in the control and annual sites

1. Wrangle the urchins data to only include Naples Reef observations for purple urchins observed in 2010. 
2. Use an exploratory faceted histograms and QQ-plot for the size distribution of purple urchins at Naples in 2010, separated by treatment (Control and Annual), to assess normality. 
3. Create a finalized basic summary table (with table caption above) of descriptive statistics for both groups. 

## Insert data and clean it for purple urchin from the Naples Reef in 2010
```{r}
purple_urchins_naples_2010 <- urchins %>% # create a new data class for NAPL in 2010
  clean_names() %>% 
  filter(common_name == "Purple Urchin") %>% # filter for Purple Urchin species only
  filter(site == "NAPL") %>% # filter for Naples Reef only
  mutate(mdy = mdy(date)) %>% # mdy() will convert the dates to a date class
  mutate(year = year(mdy)) %>% 
  filter(year == 2010) %>% # filter only show data for 2010 
  mutate(treatment_type = case_when(
    treatment == "ANNUAL" ~ "Experimental", # renaming annual as experimental so it's label is more intuitive
  treatment == "CONTROL" ~ "Control"))
```

## Let's compare the control and annual data for this dataset with a histogram

```{r}
# use ggplot to create an exploratory histogram that compares purple urchin size in the 2 observation types
ggplot(data = purple_urchins_naples_2010,
       aes(x = size)) + 
  geom_histogram(aes(fill = treatment_type, color = treatment_type), bins = 10) + 
  facet_wrap(~ treatment_type) +
  theme_minimal() +
  scale_color_manual(values = c("white", "white")) +
  theme(legend.position = "none") + # I don't feel like there's a need for a legend in this graph because it already has a label 
  labs(title = "Distribution of purple urchin sizes in annual and control sites",
       caption = "Fig 1. Observations of purple urchin sizes at Naples Reef 2010 annual and control sites",
       x = "Size (cm)",
       y = "Count") +
  theme(plot.title = element_text(hjust = 0.5))
```

## Now let's assess how normally distributed this data is with a qq-plot
```{r}
# qqplot will compare the values in the distributions to a theoretical normally distributed dataset to assess normality
ggplot(data = purple_urchins_naples_2010, aes(sample = size)) +
  geom_qq(aes(sample = size, color = treatment_type)) + 
  # geom_qq() doesn't use an "x" variable, instead it uses "sample" because those are the values that will be compared to the normal data
  stat_qq_line() +
  theme_minimal() +
  facet_wrap(~ treatment_type)
```

These two exploratory graphs suggest that the distribution of purple urchin sizes in control and experimental sites are farily normally distributed, according to the qq-plot's fairly linear trend. The control sites seem to have a slightly more normal distribution whereas experimental sites have a very strong trend for sizes around 3 cm. 

## Now make a table to summarize and compare the data
```{r}
purple_urchins_naples_2010_summary <- purple_urchins_naples_2010 %>% 
  group_by(treatment_type) %>% 
  summarize(mean_size = mean(size, na.rm = TRUE),
            median_size = median(size, na.rm = TRUE),
            sd_size = sd(size, na.rm = TRUE),
            sample_size = n())

# Create a table to present the data using kableExtra()
purple_urchins_naples_2010_summary %>% 
  kable(digits = c(2, 2, 2, 2),
        col.names = c("Common Name",
                      "Mean Size (cm)",
                      "Median Size (cm)",
                      "Standard Deviation (cm)",
                      "Sample Size"),
        caption = "<center>Table 1. Summary of Statistics of Purple Urchin Size Distribution at Naples Reef in 2010<center>") %>% 
  kable_styling(bootstrap_options = "striped", full_width = FALSE) # I just like how this bootstrap looks... 
```
