---
title: "a3_task3_shayan_kaveh"
author: "Shayan Kaveh"
date: "10/23/2021"
output: html_document
---

## Overview

This code explores data on red and purple sea urchins recorded by Santa Barbara Coastal Long Term Ecological Research program divers.

**Data:** Reed, D. 2020. SBC LTER: Reef: Long-term experiment: Kelp removal: Urchin size frequency distribution ver 20. Environmental Data Initiative. <https://doi.org/10.6073/pasta/5a1e9ef03aa47bd2225c0bb98a02a63b>.

## Set Up
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(here)
library(janitor)
library(lubridate)
```

## Read in data
```{r}
urchins <- read_csv(here("data", "urchins.csv")) %>% 
  clean_names()
```

## Section 1
1. Data visualization: compare the distribution of sizes of purple and red urchins with a histogram

```{r}
ggplot(data = urchins, 
         aes(x = size)) + 
  geom_histogram(bins = 10, 
                 aes(fill = common_name)) +
  theme_minimal() +
  facet_wrap(~ common_name) +
  labs(title = "Distribution of Purple and Red Urchin Sizes",
       key = "Common Name",
       x = "Size (cm)",
       y = "Count",
       caption = "Fig 1. Distribution of urchin sizes based on color from Santa Barbara Coastal Long Term Ecological Research",
       fill = "Common Name") + # "fill" edits the caption name
  theme(plot.title = element_text(hjust = 0.5, size = 15),
        plot.caption = element_text(hjust = 0.05))
```

