---
title: "a3_task3_shayan_kaveh"
author: "Shayan Kaveh"
date: "10/23/2021"
output: html_document
---

## Overview

This code explores data on red and purple sea urchins recorded by Santa Barbara Coastal Long Term Ecological Research program divers. The research was part of a giant kelp removal project that examines how urchin population and sizes are affected by the removals. This code looks at the distribution and size of red and purple urchins at 5 sites from 2008 to 2020, which we narrow down to look at the change of purple urchin population. 

**Data:** Reed, D. 2020. SBC LTER: Reef: Long-term experiment: Kelp removal: Urchin size frequency distribution ver 20. Environmental Data Initiative. <https://doi.org/10.6073/pasta/5a1e9ef03aa47bd2225c0bb98a02a63b>.

## Set Up

```{r setup, include= TRUE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(here)
library(janitor)
library(lubridate) # this allows us to create a date class
library(kableExtra) # use this to make pretty tables
library(RColorBrewer) # installed this color palette set to have nice graph color sets more easily!
```

## Read in data

```{r}
urchins <- read_csv(here("data", "urchins.csv")) %>% 
  clean_names() # make it lowersnake case 
```

## Section 1

1.  Data visualization: compare the distribution of sizes of purple and red urchins with a histogram

```{r}
# create a histogram in ggplot2 to compare sizes of purple and red urchins in all of the data
ggplot(data = urchins, 
         aes(x = size)) + # histogram only uses x variable~
  geom_histogram(bins = 12, # bins auto sets to 30, so made it smaller to show the trend better
                 aes(fill = common_name, color = common_name)) + 
  theme_gray() +
  scale_fill_manual(values = c("mediumpurple1", "indianred1")) + # scale_fill_manual() allows me to assign specific colors to each facet based on the vector
  scale_color_manual(values = c("white","white")) +
  facet_wrap(~ common_name) + # facet wrapping by common name to show the differences between the two urchin species
  labs(title = "Distribution of Purple and Red Urchin Sizes",
       key = "Common Name",
       x = "Size (cm)",
       y = "Count",
       caption = "Fig 1. Distribution of urchin sizes based on color from Santa Barbara Coastal Long Term Ecological Research",
       fill = "Common Name") + # "fill" edits the caption name
  theme(plot.title = element_text(hjust = 0.5, size = 15), # hjust allows the title to be centered
        plot.caption = element_text(hjust = 0.05),
        legend.position = "none")
```

2.  Descriptive Statistics Table: Create a finalized summary table that contains the following statistics for purple and red urchins (still ignoring site as a variable, but including species as a grouping variable): mean, median, standard deviation, and sample size. Use kableExtra (or another package for tables in R) to finalize your table for your knitted document. Add a table caption above your table

```{r}
urchin_summary <- urchins %>% 
  group_by(common_name) %>% # group by common name to separate the two urchin species
  summarize(mean_size = mean(size, na.rm = TRUE), # can also use round() here, but that would make me lose some data in my data frame, which I don't want to. in this case, I'll use define digits in kable() so that values only round for the output table
            median_size = median(size, na.rm = TRUE), # this will create a column that shows the median size of the urchins
            sd_size = sd(size, na.rm = TRUE), #sd() will show the standard deviation of the data
            sample_size = n()) # n() gives the sample size

# Create a table to present the data using kableExtra()
urchin_summary %>% 
  kable(digits = 2, # in kable() can use digits = x to round the values in the table - can use a c("x","y") to define the number of digits for each column
        col.names = c("Common Name",
                      "Mean Size (cm)",
                      "Median Size (cm)",
                      "Standard Deviation (cm)",
                      "Sample Size"),
        caption = "<center>Table 1. Summary of Statistics of Purple and Red Urchins<center>") %>%  # <center> will center the text 
     kable_styling(bootstrap_options = "striped", full_width = FALSE, position = "center") # bootstrap will make the table look all nice with different colors, I like it
 
```
With far over 30 samples, the data does adhere to the central limit theorem; although the purple urchin distribution is slightly right-skewed. As can be seen by the smaller standard deviation and difference in mean and median, the purple urchin data has more concentrated count with less spread. It is observed that red urchin on average are larger than purple by 2.3 cm or a percent difference of 41%. 

## Section 2
Explore purple urchin counts by site & year

```{r}
# Create a subset for purple urchins, create a column to show the site year, and write out the full names of the sites

purple_urchins <- urchins %>% 
  filter(common_name == "Purple Urchin") %>% # filter to only include purple urchins
  mutate(mdy = mdy(date)) %>% # create a date class column
  mutate(year = year(mdy)) %>% # create a column with just the year
  mutate(site_full = case_when( # case_when() allows us to say when we see this data value, we want to change it to this
    site == "NAPL" ~ "Naples Reef",
    site == "AQUE" ~ "Arroyo Quemado Reef",
    site == "CARP" ~ "Carpinteria Reef",
    site == "MOHK" ~ "Mohawk Reef",
    site == "IVEE" ~ "Isla Vista Reef"
  )) 

# create a table to see the count of purple urchins in each site per year
purple_count <- purple_urchins %>% 
  group_by(site_full, year) %>% 
  summarize(count = n())
```

Now let's make a graph to visualize the urchin count at each site over the years
```{r}
ggplot(data = purple_count,
       aes(x = year, y = count)) +
  geom_col(aes(fill = site_full, color = site_full)) + # chose to use a column graph because it shows the trend in a very visible, discrete way
  facet_wrap(~ site_full) +
  # if I wanted to show the axes on each faceted graph, I could add facet_wrap(scales = 'free_x') and the x-axis would appear on each graph
  theme_gray() +
  scale_color_manual(values = c("white", "white", "white", "white", "white")) +
  scale_fill_brewer(palette = "Set2") +
  # installed the brewer package to use the pre-created palettes they have~
  labs(fill = "Site",
       x = "Year",
       y = "Total Number of Purple Urchins",
       title = "Purple Urchin Counts from 2008 - 2020",
       caption = "Fig 2. Purple urchin counts at each observation site from 2008 to 2020 based on \n Santa Barbara Coastal Long Term Ecological Research") +
   theme(plot.title = element_text(hjust = 0.5, size = 15),  
        plot.caption = element_text(hjust = 1),
        legend.position = "none" # removed the legend because the graph already shows the title of the reefs already
        )
```

Considering these sample populations were recorded at long-term kelp removal experiment sites, this data may not accurately represent the overall purple urchin population. Although there is a trend that populations were largest at the start of the data collection, we need to know how the habitat changed and whether these sample population habitats accurately represent the overall Santa Barbara coast. The large drop in counts during 2020 could also be a result of changes to sample collection due to the COVID-19 pandemic. It's also important to note that this data combines results from control and experimental sites, so there is no way to differentiate the overall trend from what might be experimental results. 
